Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-03-20 doc/network.tex
Trying to simplify

An infected individual may contact the same susceptible individual multiple times, but only the first effective contact gives rise to infection in a given individual (after this, they are no longer susceptible).
Therefore, we expect realized generation intervals to be shorter than intrinsic generation intervals, on average, in a limited contact network.
To explore the effect of multiple contacts on realized generation intervals, we first consider the infection process from an ``egocentric'' point of view, taking into account infectious contacts made by a single infector.
We define the egocentric kernel as the rate at which secondary infections are realized by a single primary case $a$ in the absence of other infectors:
\begin{equation}
\hat{K}_a(\tau) = K_a(\tau) \exp \left(- \delta_a \int_0^\tau K_a(s) ds\right),
\end{equation}
where $K_a(\tau)$ is the individual-level intrinsic kernel and $e^{- \delta_a \int_0^\tau K_a(s) ds}$ is the probability that a susceptible acquaintance has not yet been contacted by individual $a$.
The dilution term, $\delta_a$, models how contacts are distributed among susceptible acquaintances.
For example, when infectious contacts are distributed equally in a homogeneous population, we have $\delta_a = 1/(N-1)$, where $N$ is the population size.

@@

For example, consider a susceptible-exposed-infected-recovered (SEIR) model.
This widely used model assumes that latent and infectious periods are exponentially distributed.
The intrinsic generation-interval distribution that corresponds to this model can be written has \citep{svensson2015influence}:
\begin{equation}
g(\tau) = \frac{\sigma \gamma}{\sigma - \gamma} \left(e^{-\gamma t} - e^{-\sigma t}\right),
\end{equation}
where $1/\sigma$ and $1/\gamma$ are mean latent and infectious periods, respectively.
Assuming that per-pair contact rate is $\lambda$ for any pair, we obtain the following egocentric generation-interval distribution is given by:
\begin{equation}
\hat{g}(\tau) = \frac{\sigma (\gamma + \lambda)}{\sigma - (\gamma + \lambda)} \left(e^{-(\gamma + \lambda)t} - e^{-\sigma t}\right)
\end{equation}
In this scenario, per-pair contact rate effectively decreases mean infectious period: an infected individual can no longer infect anyone (hence no longer infectious) when all its susceptible acquaintances have ben infected.

This calculation can be validated by simulating stochastic infection processes on a ``star'' network (i.e, a single infected individual at the center connected to multiple susceptible individuals who are not connected with each other).
Simulations confirm that in this case the distribution of contact times matches the intrinsic generation-interval distribution, while the distribution of realized generation times matches the egocentric generation-interval distribution (\fref{local}) -- generation times are shorter on average because contacts with already-infected susceptibles do not lead to new infections.

>>>>

An infected individual may contact the same susceptible individual multiple times, but only the first effective contact gives rise to infection in a given individual (after this, they are no longer susceptible).
Therefore, we expect realized generation intervals to be shorter than intrinsic generation intervals, on average, in a limited contact network.
To explore the effect of multiple contacts on realized generation intervals, we first consider the infection process from an ``egocentric'' point of view, taking into account infectious contacts made by a single infector.
We define the egocentric kernel as the rate at which secondary infections are realized by a single primary case $a$ in the absence of other infectors:
\jd{We can't use dilution directly in the hazard term: it needs to be multiplied by something like big R. I would put $R\delta$ there. The extreme dilution examples would be mean degree (if highly connected people infect that much more) or individual degree (if they don't infect more at all.)}
\begin{equation}
\hat{K}_a(\tau) = K_a(\tau) \exp \left(- \delta_a \int_0^\tau K_a(s) ds\right),
\end{equation}
where $K_a(\tau)$ is the individual-level intrinsic kernel and $e^{- \delta_a \int_0^\tau K_a(s) ds}$ is the probability that a susceptible acquaintance has not yet been contacted by individual $a$.
The dilution term, $\delta_a$, models how contacts are distributed among susceptible acquaintances.
For example, when infectious contacts are distributed equally in a homogeneous population, we have $\delta_a = 1/(N-1)$, where $N$ is the population size.

@@

For example, consider a susceptible-exposed-infected-recovered (SEIR) model.
This widely used model assumes that latent and infectious periods are exponentially distributed.
The intrinsic generation-interval distribution that corresponds to this model can be written has \citep{svensson2015influence}:
\begin{equation}
g(\tau) = \frac{\sigma \gamma}{\sigma - \gamma} \left(e^{-\gamma t} - e^{-\sigma t}\right),
\end{equation}
where $1/\sigma$ and $1/\gamma$ are mean latent and infectious periods, respectively.
Assuming that per-pair contact rate is $\lambda$ for any pair, we obtain the following egocentric generation-interval distribution is given by:
\begin{equation}
\hat{g}(\tau) = \frac{\sigma (\gamma + \lambda)}{\sigma - (\gamma + \lambda)} \left(e^{-(\gamma + \lambda)t} - e^{-\sigma t}\right)
\end{equation}
In this case, with fixed infectiousness during the infection period, the effect of accounting for pairwise contacts is the same as an increase in the recovery rate: infecting a susceptible contact is analogous to no longer being infectious (since the contact cannot be infected again).

This calculation can be validated by simulating stochastic infection processes on a ``star'' network (i.e, a single infected individual at the center connected to multiple susceptible individuals who are not connected with each other).
Simulations confirm that in this case the distribution of \emph{contact} times matches the intrinsic generation-interval distribution, while the distribution of realized generation times (i.e., \emph{infection} times) matches the egocentric generation-interval distribution (\fref{local}).
